---
title: "reproducible_research_assgn1"
output:
  html_document: default
  pdf_document: default
---

# Quantified Self
## Monitoring the number of steps for an individual in the months of October and November, 2012

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This document will delve into the activity.csv dataset and will cover;  

1. Reading in the data
2. Histogram of the total number of steps taken each day
3. Mean and median number of steps taken each day
4. Time series plot of the average number of steps taken        
5. The 5-minute interval that, on average, contains the maximum number of steps
6. Code to describe and show a strategy for imputing missing data
7. Histogram of the total number of steps taken each day after missing values are imputed
8. Panel plot comparing the average number of steps taken per 5 minute interval across weekdays and weekends

I will address each one by one.

### 1. Reading in the Data


```{r}
setwd("C:/Users/eo6405/Desktop/Misc - R/Coursera/Reproducible Research/Assignment 1")

library(readr)
activity <- read.csv("repdata%2Fdata%2Factivity/activity.csv")
summary(activity)
```


### 2. Histogram of the total number of steps taken each day

```{r}
library(ggplot2)

ggplot(activity, aes(date, steps, group = 1)) + 
        stat_summary(geom = "bar", fun.y = "sum", fill = "lightblue") + 
        labs(x = "Date", y = "Total number of Steps", title = "Total Steps per Day") +
        theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

### 3. Mean and Median Steps per Day

```{r}
library(dplyr)

activity %>% 
        group_by(date) %>%
        summarise(median(steps), mean(steps)) -> activity_by_day

names(activity_by_day) <- c("Date", "Median", "Mean")

activity_by_day

```

Now plotting this...
```{r}
ggplot(activity, aes(date, steps, group = 1)) +
        stat_summary(aes(date, steps, color = "Mean"), geom = 'line', fun.y = "mean") +
        stat_summary(aes(date, steps, color = "Median"), geom = 'line', fun.y = "median") +
        theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
        labs(x = "Date", y = "", title = "Mean and Median Steps") +
        scale_color_manual(name = "", values = c(Mean = "blue", Median = "red"))
```


Note that the median for each day appears to be 0.

### 4. Timeseries plot of average number of steps taken
```{r}
ggplot(activity, aes(interval, steps, group = 1)) +
        stat_summary(aes(interval, steps), geom = 'line', fun.y = "mean", color = "blue") +
        theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
        labs(x = "Interval", y = "Steps", title = "Time Series : Avg Steps")
```

### 5. 5 Minute interval that contains largest number of Steps


```{r}
activity_complete <- activity[complete.cases(activity),]

activity_complete %>%
        group_by(interval) %>%
        summarise(Avg = mean(steps)) %>%
        arrange(desc(Avg)) -> steps_by_interval

head(steps_by_interval)

```

Interval 835 has the highest steps on average.

### 6. Strategy for imputing data

Methodology in imputing missing data is that, for that specific interval, take the interval average. I create an 'Impute Mean' function which I then apply, by level in Interval, to Steps. This has been saved in 'activity_imputed.'

```{r}

impute.mean <- function(x) replace(x, is.na(x), mean(x, na.rm = TRUE))
activity %>%
        group_by(interval) %>%
        mutate(steps = impute.mean(steps)) -> activity_imputed

```


### 7. Histogram of total number of steps taken each day, with imputations.

This will be the same plot as in 2.

```{r}
ggplot(activity_imputed, aes(date, steps, group = 1)) + 
        stat_summary(geom = "bar", fun.y = "sum", fill = "lightblue") + 
        labs(x = "Date", y = "Total number of Steps", title = "Total Steps per Day - All Data") +
        theme(axis.text.x = element_text(angle = 90, hjust = 1))
```


### 8. Panel plot comparing the average number of steps taken per 5 minute interval across weekdays and weekends

To do this, I;

1. Create a variable which pulls the day of the week out
2. Create a vector which names all 'weekdays'
3. Create a factor where, if the weekday falls within 'weekdays', it is named a 'weekday.' If not, it is a 'weekend.'
4. Plot on two separate axes

```{r}
activity_imputed$day <- weekdays(as.Date(activity_imputed$date))

weekday <- c('Monday', 'Tuesday', 'Wednesday', 'Thursday', 'Friday')

activity_imputed$weekday <- factor((activity_imputed$day) %in% weekday, 
                   levels=c(FALSE, TRUE), labels=c('weekend', 'weekday'))

ggplot(activity_imputed, aes(interval, steps)) +
        stat_summary(geom = "bar", fun.y = "mean", color = "red") +
        facet_grid(weekday~.) +
        labs(x = "Date", y = "Total Number of Steps", title = "Steps - Weekend vs Weekday") +
        geom_smooth(se = FALSE, size = 1, color = "black") +
        theme(axis.text.x = element_text(angle = 90, hjust = 1))
```


This concludes my assignment.



